Comparison of dimensionality reduction schemes for parallel global optimization algorithms \thanks{The study was supported by the Russian Science Foundation, project No 16-11-10150}

### 2018-04-07 2018/RSD/MultiEvolvents/RSD.tex
RSD: add new experiments

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
%%%%%ДОБАВИЛ ДЛЯ РУССКОГО ТЕКСТА
\usepackage[utf8x]{inputenc}
\usepackage[english,russian]{babel}
\usepackage{cmap}
%%%%%
\usepackage{eqnarray,amsmath}
\usepackage{subfig}
\usepackage{graphicx}
\usepackage{amsfonts}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

@@

This allows problems to be solved in which function values may not be deter-mined for the entire search domain. Under this approach, solving multidimen-sional problems is reduced (using Peano-type space-filling curves) to solving equivalent one-dimensional problems.

@@

\par
The computational scheme obtained as a result of the dimensionality reduction consists of the following
(see Figure \ref{fig:peanoCUsage}):
\begin{itemize}
  \item The optimization algorithm performs the minimization of the reduced one-dimensional
  function \(\varphi(y(x))\) from (\ref{eq:oneDimTask}),
  \item After determining the next trial point \(x\), a multidimensional image \(y\) is calculated by using the
mapping \(y(x)\),
  \item The value of the initial multidimensional function \(\varphi(y)\) is calculated at the point \(y\in D\),
  \item The calculated value \(z=\varphi(y)\) is used further as the value of the reduced one-dimensional function \(\varphi(y(x))\) at the point \(x\).
\end{itemize}

\begin{figure}
    \centering
    \includegraphics[width=0.55\textwidth]{pictures/peanoCUsage.eps}
    \caption{The computational scheme for obtaining the value of the reduced one-dimensional function \(\varphi(y(x))\)}
    \label{fig:peanoCUsage}
\end{figure}

@@

In Fig.~\ref{6_fig_9} two evolvents being the approximations to Peano curves for the case $N=2$ are presented as an illustration. The maximum number of various rotations of the evolvents mapping the $N$-dimensional hypercube onto a one-dimensional interval is $2^N$. The employment of all possible rotations might appear to be redundant. In this case, one can select only a part of all rotations.

\begin{figure}[t]
  \centering
  \includegraphics[width=0.8\linewidth]{pictures/rotated_1.pdf}
  \caption{Two rotated evolvents on the same plane}
  \label{6_fig_9}
\end{figure}

As a possible approach, one can propose to generate the new evolvents by the rotation of the initial evolvents on the angle of $\pm\pi/2$ in each of the coordinate planes. For example, the rotation matrices in the plane $(y_2,y_4)$ for  $N=5$ are presented below.
\begin{eqnarray*}
 \begin{pmatrix}
  1 & 0 & 0 & 0 & 0 \\
  0 & 0 & 0 & -1 & 0 \\
  0 & 0 & 1 & 0 & 0 \\
  0 & 1 & 0 & 0 & 0 \\
  0 & 0 & 0 & 0 & 1
 \end{pmatrix}
\qquad \qquad
 \begin{pmatrix}
  1 & 0 & 0 & 0 & 0 \\
  0 & 0 & 0 & 1 & 0 \\
  0 & 0 & 1 & 0 & 0 \\
  0 & -1 & 0 & 0 & 0 \\
  0 & 0 & 0 & 0 & 1
 \end{pmatrix}
\end{eqnarray*}

The number of such pairs of rotations is determined by the number of the coordinate planes in the space, which equals to  $N(N-1)/2$. The total number of the transformations equals to $N(N-1)$. Taking into account the initial mapping, one can conclude that this method allows to build up to $N(N-1)+1$ evolvents for mapping the $N$-dimensional domain onto the corresponding one-dimensional intervals. Moreover, the additional constraint  $g_0(y) \leq 0$ with $g_0(y)$ from (\ref{6_g0}), which arises in shifted evolvents, is absent. This method for building a set of mappings can be ``scaled'' easily to obtain more evolvents (up to $2^N$) if necessary.

@@

Кривая типа Пеано, используемая в (\ref{eq:oneDimTask}) для редукции размерности, определяется через предельный переход,
поэтому не может быть вычислена непосредтвенно. При численной оптимизации используется некоторое её приближение, являющееся
инъективной кусочно-линейной кривой. В \cite{strongin1978} было предложено неинъективное отображение сетки на отрезка $[0,1]$ на сетку в гиперкубе $D$. Каждый многомерный узел может иметь до $2^N$ одномерных прообразов. На рис. \ref{fig:noninjective} изображена кривая с самопересечениями, полученная при соединении узлов грубой многомерной сетки в порядке следования их прообразов из отрезка $2^N$, а также отмечана точка, имеющая 3 прообраза.

\begin{figure}[t]
  \centering
  \includegraphics[width=0.6\linewidth]{pictures/noninjective.pdf}
  \caption{Numerical non injective evolvent}
  \label{fig:noninjective}
\end{figure}

@@

@@

\subsection{Параллельные множественные отображения}
\label{sec:parallel_evolvents}
The reduction of the multidimensional problems to the one-dimensional ones using evolvents has such important properties as the continuity and preservation of boundedness of function divided differences. However, a partial loss of in-formation on the nearness of the points in the multidimensional space takes place since a point \mbox{\(x\in[0,1]\)} has only the left and the right neighbors while the corresponding point  \mbox{\(y(x)\in R^N\)} has the neighbors in \(2N\) directions. As a result, when using the mappings like Peano curve the images \(y',y''\), which are close to each other in the N-dimensional space can correspond to the preimages \(x',x''\), which can be far away from each other in the interval [0,1]. This property results in the excess computations since several limit points \(x',x''\) of the trial sequence generated by the index method in the interval [0,1] can correspond to a single limit point y in the \(N\)-dimensional space.
\par
One of the possible ways to overcome this disadvantage consists in using the multiple mapping \(Y^S (x)={y^1 (x),…,y^S (x)}\) instead of single evolvent \(y(x)\). To construct the set \(Y^S (x)\) different approaches can be used. For example, in \cite{strSergGO} a scheme was implemented, according to which each evolvent \(y^i (x)\) from \(Y^S (x)\) is constructed as a result of shifting the original evolvent \(y^0 (x)\) along the main diagonal of the hypercube \(D\). The set of Peano curves thus constructed allows one to obtain \(y',y''\) from \(D\) for any close multidimensional images, which differ only in one coordinate, close preimages \(x',x''\) from the interval [0,1] for the evolvent \(y^s (x),1≤s≤S\).
Using the multiple mapping allows solving initial problem (1) by parallel solving the problems
\par

\[
\min\{\varphi(y^s(x)):x\in [0,1]\}, 1≤s≤S
\]

on a set of intervals [0,1] by the index method. Each one-dimensional problem is solved on a separate processor. The trial results at the point \(x^k\) obtained for the problem being solved by particular processor are interpreted as the results of the trials in the rest problems (in the corresponding points \(x^(k_1 ),…,x^(k_S ))\). In this approach, a trial at the point \(x^k \in [0,1]\) executed in the framework of the \(s\)-th problem, consists in the following sequence of operations.
\par
1. Determine the image \(y^k=y^s (x^k)\) for the evolvent \(y^s (x)\).
\par
2. Inform the rest of processors about the start of the trial execution at the point\( y^k\) (the blocking of the point \(y^k\) ).
\par
3. Determine the preimages \(x{}^{k_s}  \in [0,1], 1≤s≤S\), of the point \(y^k\) and interpret the trial executed at the point \(y^k \in D \) as the execution of the trials in the \(S\) points \(x{}^{k_1} ,…,x{}^{k_s} \)
\par
4. Inform the rest of processors about the trial results at the point \(y^k\).
\par

The decision rules for the proposed parallel algorithm, in general, are the same as the rules of the sequential algorithm (except the method of the trial execution). Each processor has its own copy of the software realizing the computations of the problem functions and the decision rule of the index algorithm. For the organization of the interactions among the processors, the queues are created on each processor, where the processors store the information on the executed iterations in the form of the tuples: the processor number \(s\), the trial point \(x{}^{k_s}\).
\par
The proposed parallelization scheme was implemented with the use of MPI technology. Main features of implementation consist in the following. A separate MPI-process is created for each of \(S\) one-dimensional problems being solved, usually, one process per one processor employed. Each process can use p threads, usually one thread per an accessible core.
\par
At every iteration of the method, the process with the index \(s,0≤ s< S\) performs p trials in parallel at the points \(x^(s+iS),0≤i<p\). At that, each process stores all \(S_p\) points, and an attribute indicating whether this point is blocked by another process or not is stored for each point. Let us remind that the point is blocked if the process starts the execution of a trial at this point.
\par
At every iteration of the algorithm, operating within the \(s\)-th process, determines the coordinates of p «its own» trial points. Then, the interchange of the coordinates of images of the trial points \(y^(s+iS),0≤i<p, 0≤ s< S\) is performed (from each process to each one). After that, the preimages \(x^(q+iS),0≤ q<S,q≠s\) of the points received by the \(s\)-th process from the neighbor ones are determined with the use of the evolvent \(y^s (x)\). The points blocked within the \(s\)-th process will correspond to the preimages obtained. Then, each process performs the trials at the non-blocked points, the computations are performed in parallel using OpenMP. The results of the executed trials (the index of the point, the computed values of the problem functions, and the attribute of unblocking of this point) are transferred to all rest processes. All the points are added to the search information database, and the transition to the next iteration is performed.
\par

\section{Results of Numerical Experiments}
s
\subsection{Сравнение последовательных сдвиговых и вращаемых разверток}
s
\subsection{Параллельные вращаемые развертки}
s
\subsection{Параллельные вращаемые развертки для задач большой размерности}
s
\section{Вывод о целесообразности применения того или иного вида разверток для того или иного вида задач}
s

@@

\bibitem{Gergel2009}
Strongin, R.G., Gergel, V.P., Barkalov, K.A.: Parallel methods for global optimization problem solving. Journal of instrument engineering. \textbf{52}, 25--33 (2009) (In Russian)
\end{thebibliography}
\end{document}

>>>>

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
%%%%%ДОБАВИЛ ДЛЯ РУССКОГО ТЕКСТА
\usepackage[utf8x]{inputenc}
\usepackage[english,russian]{babel}
\usepackage{cmap}
%%%%%
\usepackage{eqnarray,amsmath}
\usepackage{subfig}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amssymb}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

@@

This allows problems to be solved in which function values may not be determined for the entire search domain. Under this approach, solving multidimen-sional problems is reduced (using Peano-type space-filling curves) to solving equivalent one-dimensional problems.

@@

\par
The computational scheme obtained as a result of the dimensionality reduction consists of the following:
%(see Figure \ref{fig:peanoCUsage}):
\begin{itemize}
  \item The optimization algorithm performs the minimization of the reduced one-dimensional
  function \(\varphi(y(x))\) from (\ref{eq:oneDimTask}),
  \item After determining the next trial point \(x\), a multidimensional image \(y\) is calculated by using the
mapping \(y(x)\),
  \item The value of the initial multidimensional function \(\varphi(y)\) is calculated at the point \(y\in D\),
  \item The calculated value \(z=\varphi(y)\) is used further as the value of the reduced one-dimensional function \(\varphi(y(x))\) at the point \(x\).
\end{itemize}

%\begin{figure}
%    \centering
%    \includegraphics[width=0.55\textwidth]{pictures/peanoCUsage.eps}
%    \caption{The computational scheme for obtaining the value of the reduced one-dimensional function \(\varphi(y(x))\)}
%    \label{fig:peanoCUsage}
%\end{figure}

@@

In Fig.~\ref{6_fig_9} two evolvents being the approximations to Peano curves for the case $N=2$ are presented as an illustration.

\begin{figure}[t]
  \centering
  \includegraphics[width=0.6\linewidth]{pictures/rotated_1.pdf}
  \caption{Two rotated evolvents on the same plane}
  \label{6_fig_9}
\end{figure}

Taking into account the initial mapping, one can conclude that current implementation of the method allows to build up to $N(N-1)+1$ evolvents for mapping the $N$-dimensional domain onto the corresponding one-dimensional intervals. Moreover, the additional constraint  $g_0(y) \leq 0$ with $g_0(y)$ from (\ref{6_g0}), which arises in shifted evolvents, is absent. This method for building a set of mappings can be ``scaled'' easily to obtain more evolvents (up to $2^N$) if necessary.

@@

Кривая типа Пеано, используемая в (\ref{eq:oneDimTask}) для редукции размерности, определяется через предельный переход,
поэтому не может быть вычислена непосредтвенно. При численной оптимизации используется некоторое её приближение, являющееся
инъективной кусочно-линейной кривой. В \cite{strongin1978} было предложено неинъективное отображение равномерной сетки на отрезка $[0,1]$ на равномерную сетку в гиперкубе $D$. Каждый многомерный узел может иметь до $2^N$ одномерных прообразов. На рис. \ref{fig:noninjective} изображена кривая с самопересечениями, полученная при соединении узлов грубой многомерной сетки в порядке следования их прообразов из отрезка $[0,1]$, а также отмечана точка, имеющая 3 прообраза.

%\begin{figure}[t]
%  \centering
%  \includegraphics[width=0.6\linewidth]{pictures/noninjective.pdf}
%  \caption{Numerical non injective evolvent}
%  \label{fig:noninjective}
%\end{figure}

@@

%\begin{figure}[t]
%  \centering
%  \includegraphics[width=0.6\linewidth]{pictures/smooth.pdf}
%  \caption{Smooth evolvent}
%  \label{fig:smooth}
%\end{figure}

\begin{figure}[ht]
    \centering
    \subfloat[Smooth evolvent]{{\includegraphics[width=.6\textwidth]{pictures/smooth.pdf}}\label{fig:smooth}}
    \subfloat[Numerical non injective evolvent]{{\includegraphics[width=.4\textwidth]{pictures/noninjective.pdf}}\label{fig:noninjective}}
    \caption{Different evolvents}
\end{figure}

@@

\subsection{Параллельные множественные отображения}
\label{sec:parallel_evolvents}
Using the multiple mapping allows solving initial problem (1) by parallel solving the problems
\[
\min\{\varphi(y^s(x)):x\in [0,1]\}, 1≤s≤S
\]
on a set of intervals [0,1] by the index method. Each one-dimensional problem is solved on a separate processor. The trial results at the point \(x^k\) obtained for the problem being solved by particular processor are interpreted as the results of the trials in the rest problems (in the corresponding points \(x^(k_1 ),…,x^(k_S ))\). In this approach, a trial at the point \(x^k \in [0,1]\) executed in the framework of the \(s\)-th problem, consists in the following sequence of operations.
\par
1. Determine the image \(y^k=y^s (x^k)\) for the evolvent \(y^s (x)\).
\par
2. Inform the rest of processors about the start of the trial execution at the point\( y^k\) (the blocking of the point \(y^k\) ).
\par
3. Determine the preimages \(x{}^{k_s}  \in [0,1], 1≤s≤S\), of the point \(y^k\) and interpret the trial executed at the point \(y^k \in D \) as the execution of the trials in the \(S\) points \(x{}^{k_1} ,…,x{}^{k_s} \)
\par
4. Inform the rest of processors about the trial results at the point \(y^k\).
\par

The decision rules for the proposed parallel algorithm, in general, are the same as the rules of the sequential algorithm (except the method of the trial execution). Each processor has its own copy of the software realizing the computations of the problem functions and the decision rule of the index algorithm. For the organization of the interactions among the processors, the queues are created on each processor, where the processors store the information on the executed iterations in the form of the tuples: the processor number \(s\), the trial point \(x{}^{k_s}\).
\par
The proposed parallelization scheme was implemented with the use of MPI technology. Main features of implementation consist in the following. A separate MPI-process is created for each of \(S\) one-dimensional problems being solved, usually, one process per one processor employed. Each process can use p threads, usually one thread per an accessible core.
\par
At every iteration of the method, the process with the index \(s,0\leqslant s< S\) performs p trials in parallel at the points \(x^{(s+i_S)},0\leqslant i<p\). At that, each process stores all \(S_p\) points, and an attribute indicating whether this point is blocked by another process or not is stored for each point. Let us remind that the point is blocked if the process starts the execution of a trial at this point.
\par
At every iteration of the algorithm, operating within the \(s\)-th process, determines the coordinates of p «its own» trial points. Then, the interchange of the coordinates of images of the trial points \(y^{(s+i_S)},0\leqslant i<p, 0\leqslant s< S\) is performed (from each process to each one). After that, the preimages \(x^{(q+i_S)},0\leqslant q<S,q\not=s\) of the points received by the \(s\)-th process from the neighbor ones are determined with the use of the evolvent \(y^s (x)\). The points blocked within the \(s\)-th process will correspond to the preimages obtained. Then, each process performs the trials at the non-blocked points, the computations are performed in parallel using OpenMP. The results of the executed trials (the index of the point, the computed values of the problem functions, and the attribute of unblocking of this point) are transferred to all rest processes. All the points are added to the search information database, and the transition to the next iteration is performed.
\par

\section{Results of Numerical Experiments}
The computational experiments have been carried out on the Lobachevsky supercomputer at
State University of Nizhny Novgorod. A computational node included 2 Intel
Sandy Bridge E5-2660 2.2 GHz processors, 64 GB RAM. The CPUs had 8 cores (i. e. total 16 cores
were available per a node). Все рассматриваемые алгортимы и развёртки были реализованы на языке C++ в рамках программной системы Globalizer.
Для реализации параллелизма на одном узле использована технология OpenMP, а для параллелизма на несколько узлов --- стандарт MPI.

Сравнение алгоритмов глобальной оптимизации проведено путём оценки качества решения алгоритмами выборки задач из некоторого тестового класса.
В данной статье рассматривается тестовый класс, порождаемый генератором GKLS \cite{Gaviano2003}. Данные генератор позволяет конструировать сложные многоэкстремальные задачи различной размерности. В работе рассматриваются выборки по 100 задач классов размерности 2, 3, 4, 5. Каждый класс имеет две степени сложности --- \textit{Simple} и \textit{Hard}. Параметры генератора для рассматриваемых классов приведены в \cite{Gaviano2003}.

In order to evaluate the efficiency of an algorithm on a given set of 100 problems, we will use the
operating characteristics \cite{grishaginClass}, which are defined by a set of
points on the \((K, P)\) plane where \(K\) is the average number of search trials
conducted before satisfying the termination condition when minimizing a function
from a given class, and \(P\) is the proportion of problems solved successfully.
If at a given \(K\), the operating characteristic of a method goes higher than one
from another method, it means that at fixed search costs, the former method has a
greater probability of finding the solution. If some value of \(P\) is fixed, and the
characteristic of a method goes to the left from that of another method, the former
method requires fewer resources to achieve the same reliability.

\subsection{Сравнение последовательных разверток}
С целью понять, обладает ли какой-либо из перечисленных ранее типов развёрток существенным преимуществом над другими, были построены операционные характеристики индексного метода с различными типами развёрток на классах GKLS 2d Simple и GKLS 3d Simple. The global minimum was considered to be found if the algorithm generates a trial point $y^k$ in the $\delta$-vicinity of the global minimizer, i.e. $\left\|y^k-y^\ast\right\|_\infty\leq\delta$. The size of the vicinity was selected as $\delta = 0.01\left\|b-a\right\|$. In case of GKLS $\delta=0.01$.

Во всех экспериментах параметр плотности построения развёрток $m=12$. Минимальное значение пераметра надёжности \(r\) было найдено для каждого типа развёртки перебором по равномерной сетке с шагом \(0.1\).

На классе GKLS 2d Simple при минимальном \(r\) неинъективная и гладкая развёртка обеспечивают более быструю сходимость (рис. \ref{fig:gkls2d_opt}). То же самое, наблюдается и при \(r=5.0\) (рис. \ref{fig:gkls2d_acc}). В последнем случае сдвиговая и вращаемая развёртки начинают отставать от остальных, т.к. значение \(r=5.0\) является завышенным для них.
\begin{figure}[ht]
    \centering
    \subfloat[$r=5.0$]{{\includegraphics[width=.5\textwidth]{pictures/gklsS2d_same_r_opt_pt_op.pdf}}\label{fig:gkls2d_acc}}
    \subfloat[Minimal $r$]{{\includegraphics[width=.5\textwidth]{pictures/gklsS2d_opt_pt_op.pdf}}\label{fig:gkls2d_opt}}
    \caption{Operating characteristics on GKLS 2d Simple class}
\end{figure}

На классе GKLS 2d Simple при минимальном \(r\) неинъективная и множественные развёртки имеют значительное преимущество над единственной развёрткой. Значение \(r=4.5\) является завышенным для вращаемой и сдвиговой развёрток.

\begin{figure}[ht]
    \centering
    \subfloat[$r=4.5$]{{\includegraphics[width=.5\textwidth]{pictures/gklsS3d_same_r_opt_pt_op.pdf}}\label{fig:gkls3d_acc}}
    \subfloat[Minimal $r$]{{\includegraphics[width=.5\textwidth]{pictures/gklsS3d_opt_pt_op.pdf}}\label{fig:gkls3d_opt}}
    \caption{Operating characteristics on GKLS 3d Simple class}
\end{figure}

\paragraph{Накладные расходы при использовании сдвиговой развёртки.}

\subsection{Параллельные вращаемые развертки}

\section{Вывод о целесообразности применения того или иного вида разверток для того или иного вида задач}
s

@@

\bibitem{Gergel2009}
Strongin, R.G., Gergel, V.P., Barkalov, K.A.: Parallel methods for global optimization problem solving. Journal of instrument engineering. \textbf{52}, 25--33 (2009) (In Russian)

\bibitem{Gaviano2003}
Gaviano, M., Kvasov, D.E, Lera, D., and Sergeyev, Ya.D.: Software for generation of classes of test functions with known local and global minima for global optimization. ACM Transactions on Mathematical Software 29(4), 469--480 (2003)

\bibitem{grishaginClass}
Grishagin, V.A.: Operating Characteristics of Some Global Search Algorithms. Problems of Statistical Optimization 7, 198--206 (1978) (In Russian)

\end{thebibliography}
\end{document}
